{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion}

### 2017-01-07 proposal/410.tex
small fixes

\begin{itemize}
	\item Busy-wait loops containing neither {\tt yield} nor {\tt xchg} (nor any other atomic instruction), such as {\tt while (!other\_thread\_ready) continue;}.
		This blurs the line between anti-pattern and concurrency bug:
		because it does not yield the CPU, a uni-processor machine must wait for the next timer tick (several milliseconds!) before making any progress;
		also, because it does not use atomic instructions, an optimizing compiler may reorder or even delete the loop's accesses.
		%
		Landslide also cannot easily identify it as similar to message-passing,
		appearing indistinguishable
		% not really? there would be dpor conflix, at least
		from a thread-local infinite computation,
		which is of course impossible to judge for halting \cite{entscheidungsproblem}.

		In such a case, Landslide will issue a bug report with the special message:
		{\em I have run a loop in [function name] an alarming number of times.
		This version of Landslide cannot distinguish between this loop being infinite versus merely undesirable.
		Please refer to the ``Synchronization (2)'' lecture.}
	\item Recursive mutex use (i.e., locking the same mutex twice in the same thread, then subsequently unlocking it twice).
		While it would not be difficult for Landslide's lock-sets to support recursive locking (using a nesting counter instead of a boolean flag),
		that would assume the corresponding mutex implementation provides the same semantics,
		which is risky business with student code. % ;)
		Furthermore, recursive locking is not an obvious solution to any of P2's challenges;
		far more often, it arises when a student's mutex tries to {\tt malloc} some internal state,
		which itself requires a mutex for safe allocation, which can lead to stack overflow and a crash.

@@

To analyze its effectiveness as a debugging tool for students,
I configured it to record a snapshot of the results of each use by the students:
which options were used (test program and time limit),
a snapshot of the student’s code, and the result of the test
(completed, timed out, bug found, or ctrl-C’ed).
I manually interpreted the snapshots to determine:
\begin{itemize}
	\item How many unique bugs the student found (discounting multiple runs that produced ``the same'' bug);
	\item How many of those bugs were deterministic versus concurrency bugs (did Landslide find the bug on the first interleaving or did it need to test multiple);
	\item How many of each category of bug the students fixed (determined when a subsequent run of Landslide on the test case failed to find the bug again).
\end{itemize}

>>>>

\begin{itemize}
	\item {\bf Busy-wait loops} containing neither {\tt yield} nor {\tt xchg} (nor any other atomic instruction), such as {\tt while (!other\_thread\_ready) continue;}.
		This blurs the line between anti-pattern and concurrency bug:
		because it does not yield the CPU, a uni-processor machine must wait for the next timer tick (several milliseconds!) before making any progress;
		also, because it does not use atomic instructions, an optimizing compiler may reorder or even delete the loop's accesses.
		%
		Landslide also cannot easily identify it as similar to message-passing,
		appearing indistinguishable
		% not really? there would be dpor conflix, at least
		from a thread-local infinite computation,
		which is of course impossible to judge for halting \cite{entscheidungsproblem}.

		In such a case, Landslide will issue a bug report with the special message:
		{\em I have run a loop in [function name] an alarming number of times.
		This version of Landslide cannot distinguish between this loop being infinite versus merely undesirable.
		Please refer to the ``Synchronization (2)'' lecture.}
	\item {\bf Recursive mutex use} (i.e., locking the same mutex twice in the same thread, then subsequently unlocking it twice).
		While it would not be difficult for Landslide's lock-sets to support recursive locking (using a nesting counter instead of a boolean flag),
		that would assume the corresponding mutex implementation provides the same semantics,
		which is risky business with student code. % ;)
		Furthermore, recursive locking is not an obvious solution to any of P2's challenges;
		far more often, it arises when a student's mutex tries to {\tt malloc} some internal state,
		which itself requires a mutex for safe allocation, which can lead to stack overflow and a crash.

@@

To analyze its effectiveness as a debugging tool for students,
I configured it to record a snapshot of the results of each use by the students:
which options were used (test program and time limit),
a snapshot of the student's code, and the result of the test
(completed, timed out, bug found, or ctrl-C'ed).
I manually interpreted the snapshots to determine:
\begin{itemize}
	\item How many unique bugs the student found (discounting multiple runs that produced ``the same'' bug);
	\item How many of those bugs were deterministic versus concurrency bugs (did Landslide find the bug on the first interleaving or did it need to test multiple);
	\item How many of each category of bug the students fixed (determined when a subsequent run of Landslide on the test case failed to find the bug again).
\end{itemize}

### 2017-01-07 proposal/related.tex
small fixes

In this chapter I provide a brief tour of the related work.

@@

Transactional memory (TM), first introduced in \cite{transactional-memory},
has received renewed attention in recent years with the announcement of Intel's Haswell architecture \cite{htm-haswell},
which promises to support hardware transactions using new {\tt x86} instructions.
Because hardware transactions can fail for any reason, not just on memory conflicts,
software TM remains relevant,
and recent works such as \cite{hybrid-htm-stm} and \cite{stm-relaxed-memory} enhance it to be nested with hardware transactions, or to be used on weak memory architectures, respectively;
which in turn produce ever more complicated concurrency models for people like me to tackle.
%
Testing approaches to TM are sparsely represented in the literature so far.
%The Weakest Reasonable Condition
\cite{specifying-verifying-tm} introduces several conditions to help specify/verify the correctness of TM {\em implementations},
but not that of client programs which use TM.
{\em Learning from Mistakes} \cite{learning-from-mistakes} is a study of the characteristics of many different types of concurrency bugs;
it found that TM could potentially fix some, but not all, of the studied bugs,
although in some cases it must be combined with other concurrency primitives,
which indicates a major need for verifying transactional code with model checking.

>>>>

This chapter provides a brief tour of the related work.

@@

Transactional memory (TM), first introduced in \cite{transactional-memory},
has received renewed attention in recent years with the announcement of Intel's Haswell architecture \cite{htm-haswell},
which promises to support hardware transactions using new x86 instructions.
Because hardware transactions can fail for any reason, not just on memory conflicts,
software TM remains relevant,
and recent works such as \cite{hybrid-htm-stm} and \cite{stm-relaxed-memory} enhance it to be nested with hardware transactions, or to be used on weak memory architectures, respectively;
which in turn produce ever more complicated concurrency models for people like me to tackle.
%
Testing approaches to TM are sparsely represented in the literature so far.
%The Weakest Reasonable Condition
\cite{specifying-verifying-tm} introduces several conditions to help specify/verify the correctness of TM implementations,
but not that of client programs which use TM.
{\em Learning from Mistakes} \cite{learning-from-mistakes} is a study of the characteristics of many different types of concurrency bugs;
it found that TM could potentially fix some, but not all, of the studied bugs,
although in some cases it must be combined with other concurrency primitives,
which indicates a major need for verifying transactional code with model checking.
